---
title: "WK4_Enrichment_Analysis"
format: html
---
1) Install and Load Library
```{r}
install.packages("tidyverse")
install.packages("ggridges")

BiocManager::install("clusterProfiler", update = FALSE, ask = FALSE)
BiocManager::install("org.Hs.eg.db", update = FALSE, ask = FALSE)
BiocManager::install("enrichplot", update = FALSE, ask = FALSE)
BiocManager::install("fgsea", update = FALSE, ask = FALSE)
```
```{r}
library(tidyverse)
library(ggridges)
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(fgsea)
```

2) Import Deseq2 Results and Prepare gene list by extracting most sig DEGs
```{r}

```

3) Create a gene ID map
- Convert gene symbols to ENSEMBLE and ENTREZIDs using bitr from the CLUSTERPROFILER Package
```{r}
#gene_map <- bitr(DEG_SIG_LIST,           #bitr stands for biological ID translation
#                 fromType = "",    #Input gene ID type
#                 toType = c("", ""), # Convert to output ID type
#                 OrgDb = )           #Organism Database
```

Overepresentation Analysis (ORA) via GO and KEGG Database

4) Use Gene Ontology (GO) database
- Most popular database that provides standardized terms to describe gene function
- Organizes genes into a structured hierarchy (ontology) of biological concepts
- Three main Categories: BP (Biological Process), MF (Molecular Function) and Cellular Component(CC)
- Allow us to do functional and pathway interpretation
```{r}
# GO_bp <- enrichGO(gene = , #Use the input list of sig gene IDs
#                   OrgDb = ,     # Use the Human gene annotation database
#                   keyType = ,     # Specifies input gene ID type
#                   ont = )               # Grab information for biological process
# 

```

5) Visualizations (Dotplots, Barplots, Enrichment MAP, Gene Network MAP)
```{r}
#Create Dotplot and barplot
```

```{r}
#Create Enrichment MAP
```

```{r}
#Gene Network MAP
```




6) Use Kyoto Encyclopedia of Genes and Genomes (KEGG) Database 
- A curated  database used to understand genes, proteins, and biological pathways
- Focuses on how genes interact in functional networks and cellular processes
- Only uses ENTREZID 
```{r}
# kegg <- enrichKEGG(
#   gene = , 
#   organism = ,
#   pvalueCutoff =     
# )



```

7) Visualizations is the same as above
```{r}
#Barplot and Dotplot
```

```{r}
#Enrichment Map
```

```{r}
#Gene Network Map
```

Gene Set Enrichment Analysis (GSEA)

8) Build ranked gene list
- Determine whether a set of genes shows consistent, coordinated changes between two biological conditions
- All genes are ranked based on a metric (Ex. Expression Changes)
- Produces an Enrichment Score (ES) for each gene set

- GSEA required a fully ranked gene list in a named numeric vector

```{r}
# 1. Prepare fold changes


# 2. Map SYMBOL -> ENSEMBL & ENTREZID

# 3. Merge mapped IDs with fold changes
# gene_list_df <- merge(XXX, data.frame(SYMBOL = names(YYY),  #Merge foldchange                                                                             with mapped IDs
#                       log2FC = YYY), by = "SYMBOL") %>%
#                 group_by(ENTREZID) %>%           #Group genes by ENTREZ ID
#                 summarise(log2FC = max(log2FC)) %>%  # For each gene, keep the highest log2                                                       fold change value
#                 ungroup()                     #Remove grouping strucutre to return to df


# 4: Arrange in proper format
```

9) GO Database for GSEA
Similar to Step 4 but now we use gseGO instead of enrichGo
```{r}

```

10) Visualizations (Barplots, GGRidges, Enrichment MAP, Gene Network MAP )
```{r}

```

